---
title: "Crypto Index"
author: "Zeyan Huang"
date: "2025-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(psych)
library(ggplot2)
library(dplyr)
```

# Data
```{r}
# Data preparation
# data from Coingecko
folder_path <- "market_cap_data/"
all_files <- list.files(folder_path)


dataset <- lapply(all_files, function(x){
  data <- read.csv(paste0(folder_path, x), header = TRUE)
  data <- data[-nrow(data), ]
  data$Date <- as.Date(data$date)
  data <- data[order(data$Date),]
  data <- data[!duplicated(data$Date), ]

  
  price <- as.numeric(data$market_cap)
  names(price) <- data$Date
  return(price)
}
  )

# Name the data list
file_names <- gsub("\\.csv", "", all_files)
names(dataset) <- file_names
```


```{r}
# Step 1: Convert each price series to a dataframe with Date + market cap
dataset_aligned <- lapply(names(dataset), function(name) {
  df <- data.frame(Date = as.Date(names(dataset[[name]])),
                   Cap = dataset[[name]])
  colnames(df)[2] <- name
  return(df)
})

# Step 2: Merge all dataframes by Date, using full outer join
data <- Reduce(function(x, y) merge(x, y, by = "Date", all = TRUE),
                      dataset_aligned)

# Step 3: set all NA value as 0
# data[is.na(data)] <- 0

# Set Date as row index
rownames(data) <- data$Date
data$Date <- NULL
```

### Static Index
```{r}
# Data prep
data_static_index <- data
```

```{r}
data_static_index[is.na(data_static_index)] <- 0
```
# Biplot
```{r}
pca_static_index <- prcomp(data_static_index, scale. = TRUE)

# # PC1 vs PC2
# biplot(pca_static_index, choices = c(1, 2), scale = 0, cex = 0, main = "Biplot PC1 vs PC2")
# 
# # PC1 vs PC3
# biplot(pca_static_index, choices = c(1, 3), scale = 0, main = "Biplot PC1 vs PC3")
# 
# # PC2 vs PC3
# biplot(pca_static_index, choices = c(2, 3), scale = 0, main = "Biplot PC2 vs PC3")
# 
# # PC3 vs PC4
# biplot(pca_static_index, choices = c(3, 4), scale = 0, main = "Biplot PC3 vs PC4")
```

```{r}
(pca_static_index$sdev[1:7])^2
sum((pca_static_index$sdev)^2)
sum((pca_static_index$sdev[1:2])^2)/sum((pca_static_index$sdev)^2)*100
```



```{r}
static_index <- as.matrix(data_static_index) %*% as.vector(pca_static_index$rotation[,1])

static_index <- as.data.frame(static_index)

static_index$Date <- as.Date(rownames(static_index))

plot(static_index$Date, static_index$V1, type = "l", xlab = "Date", ylab = "PCA Static Index")
```


```{r}
# Get the daily market capitalization
data_mc <- data.frame(RowSum = rowSums(data_static_index))

# If you want to keep rownames:
rownames(data_mc) <- rownames(data_static_index)
```

```{r}
# Combine the data_mc with static_index
data_mc_si <- merge(static_index, data_mc, by = "row.names")
rownames(data_mc_si) <- data_mc_si$Row.names
data_mc_si$Row.names <- NULL
```


```{r}
plot(data_mc_si$Date, data_mc_si$V1, type = "l", col = "blue", xlab = "Date", ylab = "Static Index")

par(new = TRUE)
plot(data_mc_si$Date, data_mc_si$RowSum, type = "l", col = "red", axes = FALSE, xlab = "", ylab = "")
axis(side = 4)  # Add right-side axis
mtext("Market Capitalization", side = 4, line = 3)  # Label for right Y-axis

# Add legend
legend("topright", legend = c("Static Index", "Market Capitalization"),
       col = c("blue", "red"), lty = 1, cex = 0.5)
```

```{r}
static_index_pc8 <- abs(as.matrix(data_static_index) %*% as.vector(rowSums(pca_static_index$rotation[,1:2])))

static_index_pc8  <- as.data.frame(static_index_pc8 )

static_index_pc8$Date <- as.Date(rownames(static_index_pc8))

# Combine the data_mc with static_index
data_mc_si_pc8 <- merge(static_index_pc8, data_mc, by = "row.names")
rownames(data_mc_si_pc8) <- data_mc_si_pc8$Row.names
data_mc_si_pc8$Row.names <- NULL
```

```{r}
plot(data_mc_si_pc8$Date, data_mc_si_pc8$V1, type = "l", col = "blue", xlab = "Date", ylab = "Static Index")

par(new = TRUE)
plot(data_mc_si_pc8$Date, data_mc_si_pc8$RowSum, type = "l", col = "red", axes = FALSE, xlab = "", ylab = "")
axis(side = 4)  # Add right-side axis
mtext("Market Capitalization", side = 4, line = 3)  # Label for right Y-axis

# Add legend
legend("topright", legend = c("Static Index", "Market Capitalization"),
       col = c("blue", "red"), lty = 1, cex = 0.5)
```

## Step 1: Setting
```{r}
# Choose of Nc: Every 90 days
Tn = 90

# Choose of a_base: Every 30 days
Tw = 30
```

## Calculating Nc for the first 90 days
```{r, warning=FALSE}
data_90_1_clean <- data %>%
  slice(1:90) %>%                     # Select the first 90 rows (time window)
  select(where(~ all(!is.na(.)))) %>% # Drop columns that contain any NA values
  select(where(~ all(. != 0))) %>%    # Drop columns that contain any 0 values
  select(where(~ {                    # Drop columns with 0 or NA standard deviation
    sd_x <- sd(., na.rm = TRUE)       # Calculate standard deviation ignoring NA
    is.finite(sd_x) && sd_x > 0       # Keep only columns with valid, non-zero SD
  })) %>%
  { .[, order(-as.numeric(.[1, ]))] } # Sort the remaining columns by the first row, descending     

head(data_90_1_clean)
```

```{r}
# Step 1: Calculate PC1 for all columns
pca_full_90_1 <- prcomp(data_90_1_clean, scale. = TRUE)

# Initialize a list to store PC1 for each subset of columns
pc1_list_90_1 <- list()
index_list_90_1 <- list()

# Loop through subsets of columns (from 1:2, 1:3, ..., 1:ncol)
for (i in 2:ncol(data_90_1_clean)) {
  subset_data <- data_90_1_clean[, 1:i]
  pca_subset <- prcomp(subset_data, scale. = TRUE)
  pc1_subset <- pca_subset$rotation[, 1]
  temp_index <- as.matrix(subset_data) %*% as.vector(pc1_subset)
  pc1_list_90_1[[i - 1]] <- pc1_subset  # Store in list (index starts at 1)
  index_list_90_1[[i - 1]] <- temp_index
}

# Output first few PC1 vectors in the list
# print(pc1_list[[1]])  # PC1 for columns 1:2
#print(pc1_list[[length(pc1_list)]])  # PC1 for columns 1:ncol
```

```{r}
# Initialize a vector to store the correlations
correlations_90_1 <- numeric(length(pc1_list_90_1) - 1)

# Index with all crypto
index_all_90_1 <- index_list_90_1[[length(index_list_90_1)]]

# Calculate the Correlation
for (i in 1: (length(index_list_90_1)-1)){
  correlations_90_1[i] <- cor(index_list_90_1[[i]], index_all_90_1)
}
```

```{r}
# Find the first index where correlation >= 0.9999
threshold <- 0.99
first_reach_index_90_1 <- which(correlations_90_1 >= threshold)[1]

# Print the index
print(paste("First correlation ≥ 0.99 at subset index:", first_reach_index_90_1))
```


```{r}
plot(correlations_90_1, type = "l", col = "blue", lwd = 2,
     xlab = "Subset Index", 
     ylab = "Correlation with Last Vector",
     main = "Correlation of Each Subset PC1 with Full PC1")

# Add points to the line
points(correlations_90_1, col = "red", pch = 19, cex = 0.5)

# Add a horizontal line at 0.9999
abline(h = threshold, col = "green", lty = 2)

# Add a horizontal line at correlation = 1 for reference
abline(h = 1, col = "gray", lty = 2)

# Add text annotation
text(first_reach_index_90_1, correlations_90_1[first_reach_index_90_1], 
     labels = paste0("Index ", first_reach_index_90_1, "\nCorr ", round(correlations_90_1[first_reach_index_90_1], 4)),
     pos = 4, col = "purple")
```

```{r}
# Create a data frame (table) with subset index and correlations
correlation_table_90_1 <- data.frame(
  Subset_Index = 2:(length(correlations_90_1)+1),
  Correlation_with_Last = correlations_90_1
)

# View the table
print(head(correlation_table_90_1, 10))

```

## Calculating Nc for the second 90 days
```{r, warning=FALSE}
data_90_2_clean <- data %>%
  slice(31:120) %>%                     # Select the first 90 rows (time window)
  select(where(~ all(!is.na(.)))) %>% # Drop columns that contain any NA values
  select(where(~ all(. != 0))) %>%    # Drop columns that contain any 0 values
  select(where(~ {                    # Drop columns with 0 or NA standard deviation
    sd_x <- sd(., na.rm = TRUE)       # Calculate standard deviation ignoring NA
    is.finite(sd_x) && sd_x > 0       # Keep only columns with valid, non-zero SD
  })) %>%
  { .[, order(-as.numeric(.[1, ]))] } # Sort the remaining columns by the first row, descending     

head(data_90_2_clean)
```

```{r}
# Step 1: Calculate PC1 for all columns
pca_full_90_2 <- prcomp(data_90_2_clean, scale. = TRUE)
pc1_full_90_2 <- pca_full_90_2$x[, 1]  # The first principal component scores

# Initialize a list to store PC1 for each subset of columns
pc1_list_90_2 <- list()
index_list_90_2 <- list()

# Loop through subsets of columns (from 1:2, 1:3, ..., 1:ncol)
for (i in 2:ncol(data_90_2_clean)) {
  subset_data <- data_90_2_clean[, 1:i]
  pca_subset <- prcomp(subset_data, scale. = TRUE)
  pc1_subset <- pca_subset$rotation[, 1]
  temp_index <- as.matrix(subset_data) %*% as.vector(pc1_subset)
  pc1_list_90_2[[i - 1]] <- pc1_subset  # Store in list (index starts at 1)
  index_list_90_2[[i - 1]] <- temp_index
}

# Output first few PC1 vectors in the list
# print(pc1_list[[1]])  # PC1 for columns 1:2
#print(pc1_list[[length(pc1_list)]])  # PC1 for columns 1:ncol
```

```{r}
# Initialize a vector to store the correlations
correlations_90_2 <- numeric(length(pc1_list_90_2) - 1)

# Index with all crypto
index_all_90_2 <- index_list_90_2[[length(index_list_90_2)]]

# Calculate the Correlation
for (i in 1: (length(index_list_90_2)-1)){
  correlations_90_2[i] <- abs(cor(index_list_90_2[[i]], index_all_90_2))
}
```

```{r}
plot(correlations_90_2, type = "l", col = "blue", lwd = 2,
     xlab = "Subset Index", 
     ylab = "Correlation with Last Vector",
     main = "Correlation of Each Subset PC1 with Full PC1")

# Add points to the line
points(correlations_90_2, col = "red", pch = 19, cex = 0.5)

# Add a horizontal line at correlation = 1 for reference
abline(h = 1, col = "gray", lty = 2)
```

```{r}
# Create a data frame (table) with subset index and correlations
correlation_table_90_2 <- data.frame(
  Subset_Index = 2:(length(correlations_90_2)+1),
  Correlation_with_Last = correlations_90_2
)

# View the table
print(head(correlation_table_90_2, 10))

```


## Calculating Nc for the third 90 days
```{r, warning=FALSE}
data_90_3_clean <- data %>%
  slice(61:150) %>%                     # Select the first 90 rows (time window)
  select(where(~ all(!is.na(.)))) %>% # Drop columns that contain any NA values
  select(where(~ all(. != 0))) %>%    # Drop columns that contain any 0 values
  select(where(~ {                    # Drop columns with 0 or NA standard deviation
    sd_x <- sd(., na.rm = TRUE)       # Calculate standard deviation ignoring NA
    is.finite(sd_x) && sd_x > 0       # Keep only columns with valid, non-zero SD
  })) %>%
  { .[, order(-as.numeric(.[1, ]))] } # Sort the remaining columns by the first row, descending     

head(data_90_3_clean)
```

```{r}
# Step 1: Calculate PC1 for all columns
pca_full_90_3 <- prcomp(data_90_3_clean, scale. = TRUE)
pc1_full_90_3 <- pca_full_90_3$x[, 1]  # The first principal component scores

# Initialize a list to store PC1 for each subset of columns
pc1_list_90_3 <- list()
index_list_90_3 <- list()

# Loop through subsets of columns (from 1:2, 1:3, ..., 1:ncol)
for (i in 2:ncol(data_90_3_clean)) {
  subset_data <- data_90_3_clean[, 1:i]
  pca_subset <- prcomp(subset_data, scale. = TRUE)
  pc1_subset <- pca_subset$rotation[, 1]
  temp_index <- as.matrix(subset_data) %*% as.vector(pc1_subset)
  pc1_list_90_3[[i - 1]] <- pc1_subset  # Store in list (index starts at 1)
  index_list_90_3[[i - 1]] <- temp_index
}

# Output first few PC1 vectors in the list
# print(pc1_list[[1]])  # PC1 for columns 1:2
#print(pc1_list[[length(pc1_list)]])  # PC1 for columns 1:ncol
```

```{r}
# Initialize a vector to store the correlations
correlations_90_3 <- numeric(length(pc1_list_90_3) - 1)

# Index with all crypto
index_all_90_3 <- index_list_90_3[[length(index_list_90_3)]]

# Calculate the Correlation
for (i in 1: (length(index_list_90_3)-1)){
  correlations_90_3[i] <- abs(cor(index_list_90_3[[i]], index_all_90_3))
}
```

```{r}
plot(correlations_90_3, type = "l", col = "blue", lwd = 2,
     xlab = "Subset Index", 
     ylab = "Correlation with Last Vector",
     main = "Correlation of Each Subset PC1 with Full PC1")

# Add points to the line
points(correlations_90_3, col = "red", pch = 19, cex = 0.5)

# Add a horizontal line at correlation = 1 for reference
abline(h = 1, col = "gray", lty = 2)
```

```{r}
# Create a data frame (table) with subset index and correlations
correlation_table_90_3 <- data.frame(
  Subset_Index = 2:(length(correlations_90_3)+1),
  Correlation_with_Last = correlations_90_3
)

# View the table
print(head(correlation_table_90_3, 10))

```

## Calculating Nc for the forth 90 days
```{r, warning=FALSE}
data_90_4_clean <- data %>%
  slice(91:180) %>%                     # Select the first 90 rows (time window)
  select(where(~ all(!is.na(.)))) %>% # Drop columns that contain any NA values
  select(where(~ all(. != 0))) %>%    # Drop columns that contain any 0 values
  select(where(~ {                    # Drop columns with 0 or NA standard deviation
    sd_x <- sd(., na.rm = TRUE)       # Calculate standard deviation ignoring NA
    is.finite(sd_x) && sd_x > 0       # Keep only columns with valid, non-zero SD
  })) %>%
  { .[, order(-as.numeric(.[1, ]))] } # Sort the remaining columns by the first row, descending     

head(data_90_4_clean)
```

```{r}
# Step 1: Calculate PC1 for all columns
pca_full_90_4 <- prcomp(data_90_4_clean, scale. = TRUE)
pc1_full_90_4 <- pca_full_90_4$x[, 1]  # The first principal component scores

# Initialize a list to store PC1 for each subset of columns
pc1_list_90_4 <- list()
index_list_90_4 <- list()

# Loop through subsets of columns (from 1:2, 1:3, ..., 1:ncol)
for (i in 2:ncol(data_90_4_clean)) {
  subset_data <- data_90_4_clean[, 1:i]
  pca_subset <- prcomp(subset_data, scale. = TRUE)
  pc1_subset <- pca_subset$rotation[, 1]
  temp_index <- as.matrix(subset_data) %*% as.vector(pc1_subset)
  pc1_list_90_4[[i - 1]] <- pc1_subset  # Store in list (index starts at 1)
  index_list_90_4[[i - 1]] <- temp_index
}

# Output first few PC1 vectors in the list
# print(pc1_list[[1]])  # PC1 for columns 1:2
#print(pc1_list[[length(pc1_list)]])  # PC1 for columns 1:ncol
```

```{r}
# Initialize a vector to store the correlations
correlations_90_4 <- numeric(length(pc1_list_90_4) - 1)

# Index with all crypto
index_all_90_4 <- index_list_90_4[[length(index_list_90_4)]]

# Calculate the Correlation
for (i in 1: (length(index_list_90_4)-1)){
  correlations_90_4[i] <- abs(cor(index_list_90_4[[i]], index_all_90_4))
}
```

```{r}
plot(correlations_90_4, type = "l", col = "blue", lwd = 2,
     xlab = "Subset Index", 
     ylab = "Correlation with Last Vector",
     main = "Correlation of Each Subset PC1 with Full PC1")

# Add points to the line
points(correlations_90_4, col = "red", pch = 19, cex = 0.5)

# Add a horizontal line at correlation = 1 for reference
abline(h = 1, col = "gray", lty = 2)
```

```{r}
# Create a data frame (table) with subset index and correlations
correlation_table_90_4 <- data.frame(
  Subset_Index = 2:(length(correlations_90_4)+1),
  Correlation_with_Last = correlations_90_4
)

# View the table
print(head(correlation_table_90_4, 10))

```


# Nc constituents Update function
```{r}
Nc_cal <- function(dataframe, threshold = 0.99){
  # Step 1: Calculate PC1 for all columns
  pca_full_temp <- prcomp(dataframe, scale. = TRUE)

  # Initialize a list to store PC1 for each subset of columns
  pc1_list_temp <- list()
  index_list_temp <- list()
  variance_explained_list <- list()

  # Loop through subsets of columns (from 1:2, 1:3, ..., 1:ncol)
  for (i in 2:ncol(dataframe)) {
    subset_data <- dataframe[, 1:i]
    pca_subset <- prcomp(subset_data, scale. = TRUE)
    pc1_subset <- pca_subset$rotation[, 1]
    
    pc_1_v <- (pca_subset$sdev[1])^2
    pc_all_v <- sum((pca_subset$sdev)^2)
    pc_1_explained_v_pct <- pc_1_v/pc_all_v
    variance_explained_list[[i]] <- list(pc_1_v, pc_all_v, pc_1_explained_v_pct)
    
    temp_index <- as.matrix(subset_data) %*% as.vector(pc1_subset)
    pc1_list_temp[[i]] <- pc1_subset  # Store in list (index starts at 1)
    index_list_temp[[i]] <- temp_index
  }

  # Initialize a vector to store the correlations
  correlations_temp <- numeric(length(pc1_list_temp) - 1)

  # Index with all crypto
  index_all_temp <- index_list_temp[[length(index_list_temp)]]

  # Calculate the Correlation
  for (i in 2: (length(index_list_temp))){
    correlations_temp[i] <- cor(index_list_temp[[i]], index_all_temp)
  }

  # Find the first index where correlation >= threshold
  first_reach_index_temp <- which(correlations_temp >= threshold)[1]
  chosen_crypto_list <- c(colnames(dataframe)[1:first_reach_index_temp])

  output_list <- list()
  output_list[['Index']] <- index_list_temp
  output_list[['Correlation']] <- correlations_temp
  output_list[['Nc']] <- first_reach_index_temp
  output_list[['Choosen Crypto']] <- chosen_crypto_list
  output_list[['Variance']] <- variance_explained_list
  return(output_list)
}
```

# Basic dynamic index calculation function
```{r}
dynamic_index_base <- function(dataframe, Tw = 30){
  a_list <- c()
  a_base <- as.numeric(dataframe[Tw+1,])%*%as.vector(prcomp(dataframe[1:Tw,], scale.= TRUE)$rotation[,1])
  a_list <- append(a_list, 1)
  for (i in 2:Tw){
    a_t <- (as.numeric(dataframe[Tw+i,])%*%as.vector(prcomp(dataframe[i:(Tw+i-1),], scale.= TRUE)$rotation[,1]))/a_base
    a_list <- c(a_list, a_t)
    
  }
  a_t_df <- data.frame(Value = a_list, row.names = rownames(dataframe[(Tw+1):(Tw+Tw),]))
  
  return(a_t_df)
}
```
# Dynamic Index
```{r}
dynamic_index_cal <- function(dataframe, Tn = 90, Tn_gap = 30, Tw = 30, threshold = 0.99){
  
  # Nc Choosing process
  nc_result <- list()
  n_iter <- floor((nrow(dataframe) - Tn) / Tn_gap)
  for (i in 0:n_iter){
    data_90 <- dataframe %>%
    slice((1+i*30):(90+i*30)) %>%       # Select the first 90 rows (time window)
    select(where(~ all(!is.na(.)))) %>% # Drop columns that contain any NA values
    select(where(~ all(. != 0))) %>%    # Drop columns that contain any 0 values
    select(where(~ {                    # Drop columns with 0 or NA standard deviation
      sd_x <- sd(., na.rm = TRUE)       # Calculate standard deviation ignoring NA
      is.finite(sd_x) && sd_x > 0       # Keep only columns with valid, non-zero SD
    })) %>%
    { .[, order(-as.numeric(.[1, ]))] } # Sort the remaining columns by the first row, descending
    
    nc_result[[i+1]] <- Nc_cal(data_90, threshold = threshold)
  }
  
  # Dynamic Index Calculation
  dynamic_index_results <- list()
  for (i in 1:length(nc_result)){
    data_60 <- dataframe %>%
    slice((31+i*30):(90+i*30)) %>%       # Select the required 60 rows (time window)
    select(where(~ all(!is.na(.)))) %>% # Drop columns that contain any NA values
    select(where(~ all(. != 0))) %>%
    select(all_of(nc_result[[i]][["Choosen Crypto"]]))# Filter by selected crypto
    if (nrow(data_60) < 60){
      break
    }
    dynamic_index_results[[i]] <- dynamic_index_base(data_60)
  }
  
  final_results <- list(nc_result, dynamic_index_results)
  return(final_results)
}

```


```{r}
# Assuming dynamic_index_results is in second part of list
dynamic_index_results <- c[[2]]

# Remove NULLs if any
dynamic_index_results <- Filter(Negate(is.null), dynamic_index_results)

# Combine
combined_dynamic_index <- do.call(rbind, dynamic_index_results)

# View
head(combined_dynamic_index)
```
```{r}
data_mc_filtered <- data_mc[rownames(combined_dynamic_index), ]

# Step 3: Combine by columns (cbind)
combined_df <- cbind(combined_dynamic_index, data_mc_filtered)

# View result
head(combined_df)
```

```{r}
combined_df$Date <- as.Date(rownames(combined_df))
plot(combined_df$Date, 1000*combined_df$Value, type = "l", col = "blue", xlab = "Date", ylab = "Dynamic Index")

par(new = TRUE)
plot(combined_df$Date, combined_df$data_mc_filtered, type = "l", col = "red", axes = FALSE, xlab = "", ylab = "")
axis(side = 4)  # Add right-side axis
mtext("Market Capitalization", side = 4, line = 3)  # Label for right Y-axis

# Add legend
legend("topright", legend = c("Dynamic Index", "Market Capitalization"),
       col = c("blue", "red"), lty = 1, cex = 0.5)
```



# PAST CODE


```{r}
R <- cor(data)

# Drop columns with 0 or NA standard deviation
data_clean <- data[, apply(data, 2, function(x) {
  sd_x <- sd(x, na.rm = TRUE)
  is.finite(sd_x) && sd_x > 0
})]
```
```{r}
R_clean <- cor(data_clean)
```

```{r}
# Extract eigenvalues and eigenvectors
lambda_pca <- eigen(R_clean)$values
V_pca <- eigen(R_clean)$vectors

# Compute variance contribution (percentage of total variance)
variance_explained <- (lambda_pca / sum(lambda_pca)) * 100

# Create a table showing eigenvalues and variance contribution
pca_table <- data.frame(
  Component = 1:length(lambda_pca),
  Eigenvalue = lambda_pca,
  Variance_Explained = variance_explained
)

# Sort in decreasing order
pca_table <- pca_table[order(-pca_table$Variance_Explained),]

# Print results
cat("\nPCA Eigenvalues and Variance Explained:\n")
print(pca_table)
```

```{r}

```


```{r}
# Parameters
window_size <- 30
step <- window_size
n_windows <- floor(nrow(data) / (window_size))
dates <- as.Date(rownames(data))

results <- list()

for (i in 0:(n_windows - 1)) {
  start_train <- i * window_size + 1
  end_train <- start_train + window_size - 1
  if (end_train > nrow(data)) end_train <- nrow(data)
  start_test <- end_train + 1
  end_test <- start_test + window_size - 1

  if (end_test > nrow(data)) break

  train_data <- data[start_train:end_train, ]
  test_data <- data[start_test:end_test, ]
  
  # Drop columns that are all zeros in training window
  valid_cols <- colSums(train_data != 0) > 0
  train_data <- train_data[, valid_cols]
  
  
  train_data <- train_data[, apply(train_data, 2, function(x) {
  sd_x <- sd(x, na.rm = TRUE)
  is.finite(sd_x) && sd_x > 0
  })]
  test_data <- test_data[,colnames(train_data)]
  
  
  
  print(train_data)
  # Run PCA
  print(i)
  pca <- prcomp(train_data, center = TRUE, scale. = TRUE)
  pc1 <- pca$rotation[, 1]  # first principal component
  
  # Project each test day onto PC1
  projected <- as.matrix(test_data) %*% pc1


  # Store with date
  df_proj <- data.frame(
    date = dates[start_test:end_test],
    projection = as.numeric(projected)
  )
  results[[i + 1]] <- df_proj
}

# Combine all projections
daily_projections <- do.call(rbind, results)
print(head(daily_projections))
```

```{r}
ggplot(daily_projections, aes(x = date, y = projection)) +
  geom_line(color = "steelblue", size = 1) +
  labs(
    title = "Rolling PCA Projection (PC1)",
    x = "Date",
    y = "PC1 Projection Value"
  ) +
  theme_minimal() +
  theme(
    text = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(size = 14, face = "bold")
  )
```



```{r}
data_90_1 <- data[1:90, ]

# Drop the crypto which enter the market later than the windows start day
data_90_1 <- Filter(function(x) all(!is.na(x)), data_90_1)

# Drop columns with 0 or NA standard deviation
data_90_1_clean <- data_90_1[, apply(data_90_1, 2, function(x) {
  sd_x <- sd(x, na.rm = TRUE)
  is.finite(sd_x) && sd_x > 0
})]

R_90_1 <- cor(data_90_1_clean)

pca_90_1 <- prcomp(R_90_1, center = TRUE, scale. = TRUE)
pc_90_1<- data.frame(pca_90_1$rotation) # all principal component
```

```{r}
# Calculate the index contains all PC
loading_all_pc_90_1 <- rowSums(pc_90_1)
index_all_pc_90_1 <- as.matrix(data_90_1_clean) %*% loading_all_pc_90_1
```

```{r}
correlation_ls_90_1 <- list()
k <- 1
threshold <- 0.9999

while (k <= ncol(pc_90_1)) {
  # Step 1: Get the first k PCs
  pc_k <- pc_90_1[,1:k]
  # Step 2: Get the row sum
  if (k == 1) {
    loading_k_pc <- pc_k
  } else {
    loading_k_pc <- rowSums(pc_k)
  }
  # Step 3: Get the index with k PC
  index_k_pc <- as.matrix(data_90_1_clean) %*% loading_k_pc
  index_k_pc <- as.vector(index_k_pc)
  # Step 4: Calculate the Correlation between index-k-pc and index-all-pc
  corr_k <- cor(index_k_pc, index_all_pc_90_1)
  correlation_ls_90_1[[k]] <- corr_k
  
  # Step 5: Break if correlation is high enough
  if (corr_k >= threshold) {
    cat("Threshold reached at k =", k, "with correlation =", corr_k, "\n")
    break
  }

  # Next k
  k <- k + 1
}
```

## Assume we use NC = 6 since the index with 6 PC has correlation more than 0.95 with index_all
```{r}
# Use the data from 61 to 180 as first period
data_61_180 <- data[61:180, colnames(data_90_1_clean)]
```


```{r}
# Parameters
window_size <- Tw
step <- window_size
n_windows <- 90
Nc <- 6
dates <- as.Date(rownames(data_61_180))

results <- list()

for (i in 0:(n_windows - 1)) {
  start_train <- i * window_size + 1
  end_train <- start_train + window_size - 1
  if (end_train > nrow(data)) end_train <- nrow(data)

  train_data <- data_61_180[start_train:end_train, ]
  test_data <- data_61_180[end_train+1, ]
  
  train_data <- train_data[, apply(train_data, 2, function(x) {
  sd_x <- sd(x, na.rm = TRUE)
  is.finite(sd_x) && sd_x > 0
  })]
  test_data <- test_data[,colnames(train_data)]
  
  # print(train_data)
  # Run PCA
  print(i)
  pca <- prcomp(train_data, center = TRUE, scale. = TRUE)
  pc <- pca$rotation[, 1:Nc]
  loading_Nc_pc <- rowSums(pc)
  
  # Project each test day onto PC
  projected <- as.matrix(test_data) %*% loading_Nc_pc
  
  if (i == 0 || i == 30 || i == 60){
    a_base <- projected
  }
  
  index <- projected/a_base

  # Store with date
  df_proj <- data.frame(
    date = dates[end_train+1],
    projection = as.numeric(index)
  )
  results[[i + 1]] <- df_proj
}

# Combine all projections
daily_projections <- do.call(rbind, results)
print(head(daily_projections))
```






## 91-180
## Calculating Nc for the first 90 days
```{r}
data_180_1 <- data[91:180, ]

# Drop the crypto which enter the market later than the windows start day
data_180_1 <- Filter(function(x) all(!is.na(x)), data_180_1)

# Drop columns with 0 or NA standard deviation
data_180_1_clean <- data_180_1[, apply(data_180_1, 2, function(x) {
  sd_x <- sd(x, na.rm = TRUE)
  is.finite(sd_x) && sd_x > 0
})]

R_180_1 <- cor(data_180_1_clean)

pca_180_1 <- prcomp(R_180_1, center = TRUE, scale. = TRUE)
pc_180_1<- data.frame(pca_180_1$rotation) # all principal component
```

```{r}
# Calculate the index contains all PC
loading_all_pc_180_1 <- rowSums(pc_180_1)
index_all_pc_180_1 <- as.matrix(data_180_1_clean) %*% loading_all_pc_180_1
```

```{r}
correlation_ls_180_1 <- list()
k <- 1
threshold <- 0.9999

while (k <= ncol(pc_180_1)) {
  # Step 1: Get the first k PCs
  pc_k <- pc_180_1[,1:k]
  # Step 2: Get the row sum
  if (k == 1) {
    loading_k_pc <- pc_k
  } else {
    loading_k_pc <- rowSums(pc_k)
  }
  # Step 3: Get the index with k PC
  index_k_pc <- as.matrix(data_180_1_clean) %*% loading_k_pc
  index_k_pc <- as.vector(index_k_pc)
  # Step 4: Calculate the Correlation between index-k-pc and index-all-pc
  corr_k <- cor(index_k_pc, index_all_pc_180_1)
  correlation_ls_180_1[[k]] <- corr_k
  
  # Step 5: Break if correlation is high enough
  if (corr_k >= threshold) {
    cat("Threshold reached at k =", k, "with correlation =", corr_k, "\n")
    break
  }

  # Next k
  k <- k + 1
}
```



